---
title: "R Notebook"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).


```{r}
install.packages("haven")
library(haven)
library(dplyr)

ffdataset <- read_dta('/Users/eunicewu/Documents/INFO370/fragile_families_yuxing/ALLMERGED_pub.dta')
utils::View(ffdataset)

```

```{r}
baseline <- read.csv(file = '/Users/eunicewu/Documents/INFO370/project-fragile-families/data/Baseline_Features.csv')
```

```{r}
year1 <- read.csv(file = '/Users/eunicewu/Documents/INFO370/project-fragile-families/data/Year1_Features.csv')
```

```{r}
year3 <- read.csv(file = '/Users/eunicewu/Documents/INFO370/project-fragile-families/data/Year3_Features.csv')
```

```{r}
year5 <- read.csv(file = '/Users/eunicewu/Documents/INFO370/project-fragile-families/data/Year5_Features.csv')
```

```{r}
year9 <- read.csv(file = '/Users/eunicewu/Documents/INFO370/project-fragile-families/data/Year9_Features.csv')
```

```{r}
#feature selection

if(!require(mlbench)){install.packages("mlbench"); require(mlbench)} # common datasets to use
if(!require(caret)){install.packages("caret", dependencies = c("Depends", "Suggests")); require(caret)} # ML package and its dependencies. This will take awhile!
if(!require(dplyr)){install.packages("dplyr"); require(dplyr)}
set.seed(370)
install.packages("ModelMetrics")
install.packages("recipes")
install.packages("DEoptimR")
```

```{r}

#View(year9)

combine0 <- baseline %>% select(idnum, m1a4, m1a11a, m1a11b) %>% 
  filter(idnum>-1, m1a4>-1, m1a11a>-1, m1a11b>-1)

combine1 <- year1 %>% select(idnum,m2b18a,m2b18b,m2b18c,m2b18d,m2b18e,m2b18f,m2b18g,m2b18h,m2c3a,m2c3b, m2c3c,m2c3d,m2c3e,m2c3f,m2c3h,m2c3i,m2c3j) %>%    filter(m2b18a>-1,m2b18b>-1,m2b18c>-1,m2b18d>-1,m2b18e>-1,m2b18f>-1,m2b18g>-1,m2b18h>-1,m2c3a>-1,m2c3b>-1, m2c3c>-1,m2c3d>-1,m2c3e>-1,m2c3f>-1,m2c3h>-1,m2c3i>-1,m2c3j>-1)

combine3 <- year3 %>% select(idnum,m3b4a,m3b4b,m3b4c,m3b4d,m3b4e,m3b4f,m3b4g,m3b4h,m3b4i,m3b4k,m3b4l,m3b5,m3c3a,m3c3b,m3c3c,m3c3d,m3c3e,m3c3f,m3c3g,m3c3h,m3c3i,m3c3j,m3c3k,m3c3l) %>% 
  filter(m3b4a>-1,m3b4b>-1,m3b4c>-1,m3b4d>-1,m3b4e>-1,m3b4f>-1,m3b4g>-1,m3b4h>-1,m3b4i>-1,m3b4k>-1,m3b4l>-1,m3b5>-1,m3c3a>-1,m3c3b>-1,m3c3c>-1,m3c3d>-1,m3c3e>-1,m3c3f>-1,m3c3g>-1,m3c3h>-1,m3c3i>-1,m3c3j>-1,m3c3k>-1,m3c3l>-1)

combine5 <- year5 %>% select(idnum,m4b4a1,m4b4a2,m4b4a3,m4b4a4,m4b4a5,m4b4a6,m4b4a7,m4b4a8) %>% 
  filter(m4b4a1>-1,m4b4a2>-1,m4b4a3>-1,m4b4a4>-1,m4b4a5>-1,m4b4a6>-1,m4b4a7>-1,m4b4a8>-1)

combine9 <- year9 %>% select(idnum,t5c13a,t5c13b,t5c13c) %>% 
  filter(t5c13a>-1,t5c13b>-1,t5c13c>-1)

join1 <- inner_join(combine0,combine1, by="idnum")
join2 <- inner_join(join1,combine3, by="idnum")
join3 <- inner_join(join2,combine5, by="idnum")
finalCombined <- inner_join(join3,combine9, by="idnum")

# add outcome variables
finalCombined$t5c13a <- as.factor(finalCombined$t5c13a)
finalCombined$t5c13b <- as.factor(finalCombined$t5c13b)
finalCombined$t5c13c <- as.factor(finalCombined$t5c13c)

```

```{r}
## Automatic Feature Selection
# control using random forest
control <- rfeControl(functions = rfFuncs, method="cv", number=5)
results <- rfe(finalCombined[,2:53], finalCombined[,56], sizes = c(1:52), rfeControl = control) # this will take AWHILE
results
ggplot(results)

# chosen features
predictors(results)
```

*TODO*: Select the most important features (up to 10) and use only those parameters moving forward

```{r}
selected_features <- predictors(results)[1:5] 
selected_features <- predictors(results)[1:10] 
selected_features
```


Determine if which features are highly correlated

## Ranking features by importance
Rank features by importance using K-Nearest Neighbors (KNN). Plot the performance
- Use a k-fold cross-validation ("repeatedcv") with k=5 folds, repeated 3 times for control.
```{r}
control <- trainControl(method="repeatedcv", number = 5, repeats = 3)

model <- train(idnum ~., data=finalCombined, method = "knn", trControl = control)

importance <- varImp(model)

ggplot(importance)
```


